---
title: "Machine Learning Course Final Assignment"
author: "Aidar Zaripov"
date: "November 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Download and read data

Firstly, we download and read the data. 

```{r a}
setwd("C:/Users/miair/datasciencecoursera")
if(!file.exists("./machinelearning")){dir.create("./machinelearning")}
trainURL <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testURL  <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(trainURL, destfile = "./machinelearning/train.csv")
download.file(testURL , destfile = "./machinelearning/test.csv" )
train    <-read.csv("./machinelearning/train.csv")
test     <-read.csv("./machinelearning/test.csv")
```

# Include required libraries

```{r a1, message=FALSE}
library("randomForest");
library("caret");
```

# Cleaning Data

Then we remove variables with NA, missing or error values.

```{r b}
missing_rate_train <- sapply(train, function(x) sum(is.na(x) | x == "" | x == "#DIV/0!"))/nrow(train)
missing_rate_test  <- sapply(test , function(x) sum(is.na(x) | x == "" | x == "#DIV/0!"))/nrow(test)
train_raw          <- train[, missing_rate_train == 0]
test_raw           <- test [, missing_rate_test == 0]
# Remove irrelevant variables
train_raw          <- train_raw [, -c(1:7)] 
test_raw           <- test_raw  [, -c(1:7)]
```

# Splitting data into validation and train sample

Now we split our dataset into validation and train sample with proportion 30/70.

```{r c}
Val         <- createDataPartition(train_raw$classe, p = 0.7)
TrainingV   <- train_raw[  Val[[1]], ] 
ValidationV <- train_raw[ -Val[[1]], ]
```

# Data preprocessing

We center and scale predictors (data prepocessing step).

```{r d}
preProcessed               <- preProcess(TrainingV[, 1:52])

Centered_Scaled_train      <- predict(preProcessed, TrainingV[, 1:52])
Final_CS_train             <- data.frame(classe = TrainingV$classe, Centered_Scaled_train)

Centered_Scaled_validation <- predict(preProcessed, ValidationV[, 1:52])
Final_CS_validation        <- data.frame(classe = ValidationV$classe, Centered_Scaled_validation)

Centered_Scaled_test       <- predict(preProcessed, test_raw[, 1:52])
Final_CS_test              <- data.frame(classe = NA, Centered_Scaled_test)
```

# Model fitting

We fit random forest model on train sample.

```{r e}
modelRF <- randomForest(classe ~ ., data=Final_CS_train, method="class")
```

# Predictions 

Then we apply the model to get predictions on validation and test sample.

```{r f}
predictions_train         <- predict(modelRF, Final_CS_train     , type = "class")
predictions_validation    <- predict(modelRF, Final_CS_validation, type = "class")
predictions_test          <- predict(modelRF, Final_CS_test      , type = "class")
predictions_test
```

# Accuracy assesment 

Finally we get confusion matrix and calculate accuracy results.

```{r g}
confusionMatrix(predictions_train     , Final_CS_train     $classe)
confusionMatrix(predictions_validation, Final_CS_validation$classe)
```

# Conclusions

The accuracy on validation sample for the Random Forest model on our data is 99.5%.
The expected out-of-sample error is estimated as 1 - accuracy, so it is  0.5%.  

